---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# repvar

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

The goal of repvar is to find the minimum number of samples that will represent
all variables in a data set. This was built for population genetic data, but is
generalizable to any discrete data type that can be represented as an integer
matrix. 

## Installation

This package is not currently on CRAN, but you can install it like so:

```{r inst, eval = FALSE}
# install.packages("remotes") # or devtools or ghit
remotes::install_github("zkamvar/repvar")
```


## Example

Here is a basic example of how you can identify the minimum set. We will use the
pre-packaged `monilinia` data set from Everhart and Scherm 2016.

```{r example}
options(width = 120)

library("repvar")
data(monilinia)
dim(monilinia)
monilinia[1:10, 1:10]

# extract minimum set of samples
(i <- get_minimum_set(monilinia))
monilinia[i, 1:10]
```

