---
title: "Getting Started"
author: "Zhian N. Kamvar and Sydney E. Everhart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 10,
  out.width = "500px", 
  out.height = "500px"
)
```

What the repvar package does
============================

The repvar package is designed to provide a simple way of finding the minimum 
number of samples needed to **rep**resent all **var**iables in a given sparse
binary matrix. The impetus for this was driven by a need to resequence a small
set of samples from a population genetic study to serve as positive controls for
allele calls. 


How it works
------------

The main algorithm is in the function `rpv_indices()`. In it's simplest form,
the algorithm works like so:

1. count the number of observations in each column and sort them
2. set a counter to count the number of columns that haven't been accounted for
3. create a vector to store the sample names
4. while the counter is greater than zero
  A. record the first sample from each of the columns with the smallest number of observations
  B. take the union of those sample names and any other sample names that have been collected
  C. remove the columns these samples represent from the original column list
  D. set the counter to the number of columns left in the original column list
5. verify that all the columns have been accounted for by the sample names
6. return the vector of sample names

You can verify for yourself that it works. We will be using a data set 
representing 264 samples of *Monilinia fructicola* collected from peach orchards
in Georgia genotyped at 13 microsatellite loci representing 95 alleles. The data
are stored in a 264x95 matrix where each row represents a sample and each column
represents an allele. Black cells indicate the presence of an allele, blue cells
indicate an absence and white cells indicate missing data. 


```{r show_monilinia, echo = FALSE}
# Load the package and example data
library("repvar")
data("monilinia")

# Function to get midpoints of a vector
midpoints <- function(x, n) as.integer(x[-length(x)] + (x[-1] - x[-length(x)])/2)
the_loci <- c(
  'CHMFc4'  = 1,
  'CHMFc5'  = 4,
  'CHMFc12' = 10,
  'SEA'     = 13,
  'SED'     = 24,
  'SEE'     = 33,
  'SEG'     = 38,
  'SEI'     = 46,
  'SEL'     = 53,
  'SEN'     = 58,
  'SEP'     = 67,
  'SEQ'     = 75,
  'SER'     = 86
  )

image_monilinia <- function(mat, loci, col = c("powderblue", "grey10")){
  ncol  <- ncol(mat)
  ticks <- midpoints(c(loci, ncol))/ncol
  image(t(mat), axes = FALSE, col = col)
  abline(v = c((loci - 1.5)/(ncol - 1), 1 + (0.5/(ncol - 1))), lwd = 2, col = "grey97")
  axis(1, at = ticks, labels = names(loci), las = 3, tick = FALSE, padj = 0)
}

# View the data set
image_monilinia(monilinia, the_loci)
```

Immediately you should be able to see that some loci---like CHMFc4---have few
alleles that are fairly well-represented while other loci---like SEA---have
several alleles with the majority of them represented by a handful of samples.

If we run `rpv_indices()` on our data, we get a vector of samples that represent
all of the alleles:

```{r ex1}
library("repvar")
data("monilinia")
print(ids <- rpv_indices(monilinia))
length(ids)
colSums(monilinia[ids, ])
```

Let's highlight where these sampeles came from:

```{r global_show, echo = FALSE}
mon2 <- monilinia
mon2[!rownames(mon2) %in% ids, ] <- NA
image_monilinia(monilinia, the_loci)
image(t(mon2), add = TRUE, axes = FALSE, col = c("yellow", "firebrick"))
```


If we visualize just these samples, we can see that each locus is accounted for 
at least once:

```{r ex1_show, echo = FALSE, out.height = "250px", fig.height = 5}
image_monilinia(monilinia[ids, ], the_loci)
```


