---
title: "Getting Started"
author: "Zhian N. Kamvar and Sydney E. Everhart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 10,
  out.width = "500px", 
  out.height = "500px"
)
```

What the repvar package does
============================

The repvar package is designed to provide a simple way of finding the minimum 
number of samples needed to **rep**resent all **var**iables in a given sparse
binary matrix. The impetus for this was driven by a need to resequence a small
set of samples from a population genetic study to serve as positive controls for
allele calls. 


How it works
------------

The main algorithm is in the function `rpv_indices()`. In it's simplest form,
the algorithm works like so:

1. count the number of observations in each column and sort them
2. set a counter to count the number of columns that haven't been accounted for
3. create a vector to store the sample names
4. while the counter is greater than zero
     i. record the first sample from each of the columns with the smallest number of observations
     ii. take the union of those sample names and any other sample names that have been collected
     iii. remove the columns these samples represent from the original column list
     iv. set the counter to the number of columns left in the original column list
5. verify that all the columns have been accounted for by the sample names
6. return the vector of sample names

Functions in this package
-------------------------

There are four functions in this package:

1. `rpv_indices()` searches a matrix with the above algorithm and returns a
   vector of rownames for that matrix.
2. `rpv_find()` randomizes the rows of the matrix and performs `rpv_indices()`
   on the randomizations to return a list of candidates to be the minimum set.
3. `rpv_stats()` computes diversity indices for a matrix.
4. `rpv_image()` allows you to visualize your matrix as an image.

Quickstart
----------

Because there are several possibilities for minimum number of samples, the data 
should be randomized and iterated over several times. This can be done with 
`rpv_find()`:

```{r quickstart}
library("repvar")
data("monilinia")
# First, view our data
loci <- sapply(strsplit(colnames(monilinia), "[.]"), "[", 1)
rpv_image(monilinia, f = loci)

# Now randomly sample 200 times to get a set of vectors
set.seed(2018)
id_list <- rpv_find(monilinia, n = 200, cut = TRUE, progress = FALSE)
id_list
lengths(id_list)
```

We can see here that we have 3 possible sets of samples that have 30 samples
each. We can compare them using `rpv_stats()`, which calculate entropy
statistics:

```{r stats}
do.call("rbind", lapply(id_list, function(i) rpv_stats(monilinia[i, ])))
```


From this, we can see that they are all relatively equivalent, but the first set
has the least missing data and the greatest evenness (E5). 

Let's see what samples this represents:

```{r highlight}
rpv_image(monilinia, f = loci, highlight = id_list[[1]])
```

```{r highlight2, out.height = "250px", fig.height = 5}
rpv_image(monilinia[id_list[[1]], ], f = loci)
```



Detailed Example: *Monilinia fructicola*
========================================

To demonstrate why we need to shuffle the data, we will be using a data set
representing 264 samples of *Monilinia fructicola* collected from peach orchards
in Georgia genotyped at 13 microsatellite loci representing 95 alleles. The data
are stored in a 264x95 matrix where each row represents a sample and each column
represents an allele. Black cells indicate the presence of an allele, blue cells
indicate an absence and white cells indicate missing data.

```{r show_monilinia}
library("repvar")
data("monilinia")
loci <- sapply(strsplit(colnames(monilinia), "[.]"), "[", 1)
rpv_image(monilinia, f = loci)
```

Immediately you should be able to see that some loci---like CHMFc4---have few
alleles that are fairly well-represented while other loci---like SEA---have
several alleles with the majority of them represented by a handful of samples.

If we run `rpv_indices()` on our data, we get a vector of samples that represent
all of the alleles:

```{r ex1}
print(ids <- rpv_indices(monilinia))
length(ids)
colSums(monilinia[ids, ]) # all of the variables are represented
```

Let's highlight where these sampels came from:

```{r global_show}
rpv_image(monilinia, f = loci, highlight = ids)
```


If we visualize just these samples, we can see that each locus is accounted for 
at least once:

```{r ex1_show2, out.height = "250px", fig.height = 5}
rpv_image(monilinia[ids, ], f = loci)
```


However, if we shuffle the data set:

```{r ex1_shuffle}
set.seed(2018)
monilinia_shuffled <- monilinia[sample(nrow(monilinia)), ]
rpv_image(monilinia_shuffled, f = loci) # shuffled data
```

We get a different answer:

```{r ex1_show, fig.width = 20, out.width = "1000px"}
print(new_ids <- rpv_indices(monilinia_shuffled))
par(mfrow = c(1, 2))
rpv_image(monilinia, f = loci, highlight = new_ids, 
          newplot = FALSE,                 # avoid creating a new plot
          col = c("grey95", "grey10"),     # color background grey
          idcol = c("#B15928", "#FFFF99")  # reverse foreground colors
          )
rpv_image(monilinia, f = loci, highlight = ids, 
          newplot = FALSE,                 # avoid creating a new plot
          col = c("grey95", "grey10"),     # color background grey
          idcol = c("#B15928", "#FFFF99")  # reverse foreground colors
          )
par(mfrow = c(1, 1))
```

